Artificial Intelligence for the Automated Synthesis and Validation of Programs

### 2018-05-24 doc/programming-project/AI-programming.tex
small udpates to project proposal

@@

Closely related to the aims of the project is the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}. The existing algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}. To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC. In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{sergio:aprograming:icaps16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.

The computation of FSCs for generalized planning includes works that compile the generalized planning task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g. {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}). This last case requires a behavior specification of the FSC consisting on classified execution histories that (1) accept all legal execution histories leading to a goal-satisfying state, and (2) reject those that contain repeated configurations (indicating an infinite loop) and that cannot be extended (indicating a dead end)~\cite{Giacomo:FSM:ICAPS13}.

@@

\section{Metodología y Plan de Trabajo ( Apartado 5.2.ii, bases de la convocatoria)}
\label{subsec:metodologia}
Our research shows that current AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}. To illustrate this, Table~\ref{tab:programs} reports the invested time to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.

@@

\begin{enumerate}
\item {\bf T1. System design (months 1-6})
  \begin{small}
    \begin{enumerate}
    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases and the available instruction set. 
    \item Experimental design. The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks:
      \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
      \end{enumerate}
  \end{small}

@@

\item {\bf T3. Experiments and dissemination of results (months 13-24}). 
{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
\end{enumerate}

\section{Objetivos específicos del Proyecto ( Apartado 5.2.iii, bases de la convocatoria)}
\label{subsec:objectivos}
To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}. Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}. The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition under which program instructions are taken.

@@

Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception. Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition (IPC)~\cite{vallati:IPC:AI15}. This project will also help to {\bf meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks}.

>>>>

The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS), {\tt http://users.dsic.upv.es/grupos/grps/}, that is part of the {\em Department of Computer Systems and Computation} of the UPV. In the last 10 years, the GRPS has led 3 research projects of the Spanish national research plan on {\em AI planning}. Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018).

The research group for this project comprises: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.

@@

Closely related to the aims of the project is the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}. The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}. To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC. The computation of FSCs includes works that compile this task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g. {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}). This last case requires a behavior specification of the FSC consisting on classified execution histories that (1) accept all legal execution histories leading to a goal-satisfying state, and (2) reject those that contain repeated configurations (indicating an infinite loop) and that cannot be extended (indicating a dead end)~\cite{Giacomo:FSM:ICAPS13}.

@@

\section{Metodología y Plan de Trabajo ( Apartado 5.2.ii, bases de la convocatoria)}
\label{subsec:metodologia}
Our current research shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}. To illustrate this, Table~\ref{tab:programs} reports the invested time to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.

@@

\begin{enumerate}
\item {\bf T1. System design (months 1-6})
  \begin{small}
    \begin{enumerate}
    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases and the available instruction set. 
    \item Experimental design. The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks that are characterized as {\em finite state machines}.
      \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
      \end{enumerate}
  \end{small}

@@

\item {\bf T3. Experiments and dissemination of results (months 13-24}).
   \begin{small}
      \begin{enumerate}
      \item Reporting the experimental performance of our AIP approach for solving TDD programming tasks.
      \item Disseminating the obtained theoretical and empirical results by submitting research papers to the main conferences and journals in AI.        
      \end{enumerate}
\end{small}        
{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
\end{enumerate}

\section{Objetivos específicos del Proyecto ( Apartado 5.2.iii, bases de la convocatoria)}
\label{subsec:objectivos}

To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we plan to model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}. Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}. The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition under which program instructions are taken.

@@

Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception. Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}. This project will also help to {\bf meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks}.
